Four reasons your build is slow

### 2019-01-17 paper.tex
llvm section

\subsection{Bad scalability}
\label{sec:scalability}

\begin{figure}[t]
  \includegraphics[width=0.5\textwidth]{llvm-speedup-w-64}
  \caption{A comparison of the observed speedup when LLVM was built with \emph{make -j <n> for <n> from 1 to 32},
    to the potential speedups our tool predicted using the work and span calculations}
  \label{fig:llvm}
\end{figure}

Some projects may exhibit ample potential parallelism but not achieve that potential
parallelism in practice.  LLVM~\citep{LLVM:CGO04} is the largest
project we analysed, with a calculated work of approximately 12584 seconds, or 3.5 hours, and
a predicted span of 305 seconds, or 5 minutes, measured from a build of LLVM that was run without
strace. With a work of 3.5 hours and a span of 5 minutes, LLVM has a parallelism of about 41.

A parallelism of 41 lead us to believe that when built in parallel on a machine with 32 cores,
LLVM would build in about 5 minutes.  Our observations however, did not match this expectation.
In chart \ref{fig:llvm} we see two different observed speedups, indicted by the squares and
triangles.  The squares indicate the observed speedup of LLVM when run on a machine with 32 cores,
with \emph{make -j 1} to \emph{make -j 32}.  The triangles indicate the observed speedup when run
on a different machine, with 32 real cores and 32 hyperthreads, with \emph{make -j n} from where
n is most of the numbers between 1 and 32.  The dashed line indicates the upper predicted speedup,
which is the speedup the project can't exceed.  It is apparent from the chart that LLVM fails to
achieve this maximum predicted speedup.  The solid line represents the predicted lower speedup,
which is the speedup the project should achieve at minimum.  And it appears in the chart that
LLVM narrowly achieves this speedup, when built with \emph{make -j n} for most n.

The lower and upper speedup bounds are calculated using Brent's theorem \cite{brents}, which
specifies upper and lower bounds on the time, $T_P$ it will take a parallel algorithm to run on P
processors.

\begin{equation*}
  \frac{work}{P} \geq T_P \leq \frac{work}{P} + span 
\end{equation*}

>>>>

\subsection{Poor scaling}
\label{sec:scalability}

\begin{figure}[t]
  \includegraphics[width=0.5\textwidth]{llvm-speedup-w-64}
  \caption{A comparison of the observed speedup when LLVM was built
    with \emph{make -j <n> for <n> from 1 to 64}, to the potential
    speedups our tool predicted using the work and span calculations}
  \label{fig:llvm}
\end{figure}

Some projects may exhibit ample potential parallelism but not achieve
that potential parallelism in practice.  LLVM~\citep{LLVM:CGO04} is
the largest project we analysed, with a calculated work of
approximately 12584 seconds, or 3.5 hours, and a predicted span of 305
seconds, or 5 minutes (\verb|strace| instrumentation was disabled for
this measurement). With a work of 3.5 hours and a span of 5 minutes,
LLVM has a parallelism of about 41.

A parallelism of 41 lead us to believe that when built in parallel on
a machine with 32 cores, LLVM would build in not that much more than 5
minutes.  Our observations however, did not match this expectation.
In chart \ref{fig:llvm} we see two different observed speedups,
indicted by the squares and triangles.  The squares indicate the
observed speedup of LLVM when run on a machine with 32 cores, with
\emph{make -j 1} to \emph{make -j 32}.  The triangles indicate the
observed speedup when run on a different machine, with 32 cores each
with 2 hyperthreads, with \emph{make -j n} from where $n$ between 1
and 64.  The dashed line and solid lines indicates the upper and lower
bounds we would expect for parallel speedup, calculated using Brent's
bound~\cite{brents}.  It
is apparent from the chart that LLVM is consistently along the lower
bound.

This demonstrates that our automatic prediction is accurate out to a
large number of parallel jobs, but also that the LLVM build fails to
achieve as much parallel scaling as we would hope. This suggests that
although the build system is able to provide adequate parallelism, the
underlying Linux OS and \verb|make| job scheduling fail to take
advantage. To enable better scaling, potential sources of contention
such as the filesystem must be examined.
